Displacement of fishing effort by Large Scale Marine Protected Areas\footnote{Work in progress, do not circulate}

### 2018-10-19 docs/Manuscript.tex
Updated manuscript draft

  \title{Displacement of fishing effort by Large Scale Marine Protected Areas}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-16}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

\usepackage{float}
\floatplacement{figure}{H}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

@@

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environments by
2020. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018}. A significant part
of this rapid increase can be attributed to the designation of a small
number of Large Scale Marine Protected Areas (LSMPAs;
\citet{boonzaier_2016,singleton_2014,gray_2017}). These are defined as
MPAS larger than 250,000 Km\textsuperscript{2} in extension (Fig.
\ref{fig:LSMPAs_map}), and are often implemented in the pelagic
environment, where the dominant human activity is industrial fishing
\citep{toonen_2013,gray_2017,kroodsma_2018}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. Moreover, the world has seen a widespread increase in
the implementation of LSMPAs, likely as a way to rapidly achieve the
10\% target \citep{boonzaier_2016,alger_2017,sala_2018}. However, as
with customary MPAs, it is important that we understand the
socioeconomic implications of management interventions that result in
the spatial displacement of fishing effort. They \textbf{key gap in the
literature is X}. We adress it here by using new stuff and find results
that are of relevance to policy makers and advance our knowledge, which
should be incorporated when designing and planning for LSMPAs.

Due to weak property rights, limited habitat transformation, and
potentially lower management costs, pelagic MPAs provide an opportunity
to safeguard the oceans \citet{game_2009}. A growing body of literature
has shown that closing the high seas to all fishing could increase
fishery yields and profitability of fisheries, with negligible costs to
food security \citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}.

The early literature on LSMPAs focused on the inherent difficulties that
come with a pelagic (\emph{i.e.} open water) environment, where
organisms typically have large home ranges. \citet{game_2009} suggest
that most of the challenges can be overcome with the incorporation of
technology, in what then became known as Dynamic Ocean Management
\citep{maxwell_2015}. \citet{kaplan_2009} claim that very large MPAs
would result in excessive opportunity costs and that these would be
difficult to enforce. \citet{toonen_2013} show that just a small number
of LSMPAs made up 80\% of the managed areas in the ocean in 2013, and
that these can help achieve the CBD Aichi targets faster.
\citet{singleton_2014} provide an objective discussion of pros and cons
of LSMPAs.

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these human
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protectes two
thrids of the tagged grey reef sharks by effectively excluding fishing
effort. No studies have evaluated the displacement of fishing effort due
to LSMPA implementation.

The exclusion of these LSMPAs is likely to change fisher's behavior.
Theoretical models of fishing effort redistribution range from the
simplistic assumption that effort inside the bounded region disappears,
to spatially explicit models that reallocate fishing effort based on
habitat characteristics, presence of other vessels, and expected returns
\citep{smith_2003,hilborn_2006}. However, these focus on the long term
optimal equilibrium and redistribution of fishing effort may not be
optimal, especially over the first years \citep{stevenson_2013}. The few
empirical works suggest that resource users may show idiosyncratic
responses. For example, \citet{stevenson_2013} show that a network of
MPAs displaced fishing effort farther away from ports resulting in
higher \emph{perceived} costs, but that catch per unit effort also
increased. \citet{cabral_2017} analyse the redistribution of fishing and
non-fishing vessels following the implementation of MPAs in California,
and find that commercial dive boats follow a fishing-the-line pattern,
while some fishing boats follow an ideal free distribution. The way in
which fishers react to a spatial closure can have major implications in
its outcome \citep{smith_2003,hilborn_2006} highlighting the need to
understand how fishers react to the implementation of LSMPAs, and
fishing effort changes and is spatially redistributed.

The main objective of this paper is to identify behavioral responses to
the implementation of LSMPAs. We combine novel vessel tracking
technologies and causal inference techniques to identify behavioral
changes of fishing vessels due to the implementation of PIPA. We focus
on fishing hours and distance traveled as outcome variables that fishers
might adjust following implementation of a LSMPA in an impact-evaluation
fashion. Additionally, we evaluate the spatial redistribution of fishing
effort that existed within PIPA before its implementation. This work
provides novel empirical insights into fisher's responses to the
implementation of LSMPAs, and can help guide future interventions. Our
work is novel in the sense that it provides empirical evidence of the
effect of large-acale Marine Protected Areas in fishing behavior and
distribution.

The next sections are as follows: Section \ref{background} provides an
overview of the Nauru Agreement and associated countries, a description
of the fleet that operates in the region, and a brief history of PIPA.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, section \ref{discussion}
provides an extension of our results to other LSMPAs and discusses our
results.

@@

@@

The Nauru Agreement was established in 1982 by Pacific island nations to
manage their important tuna resources. PNA Members include Federated
States of Micronesia, Kiribati, Marshall Islands, Nauru, Palau, Papua
New Guinea, Solomon Islands, and Tuvalu.The Nauru Agreement regulated
access of foreign vessels (\emph{i.e.} those from non-PNA countries).
Holding \textasciitilde{}80\% of the historical purse seining grounds
within their Exclusive Economic Zones, PNA countries gained bargaining
power when providing access to foreign fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisors recomend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

@@

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. The pacific region
has historically accounted for a large portion of tuna species
production \citep{aqorau_1997}. Today, the PNA controls close to 50\% of
the global skipjack tuna production \citep{pna_website_2018}. A large
portion of these catches derive from purse seine vessels licensed under
the VDS. Fishing vessels from Australia, New Zeland, China, France,
Korea, Japan, the Philippines, Taiwan, and the United States participate
in the purse-seining VDS.

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Island Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of it was declared as no-take. In
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 Km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. The implementation of such a large
conservation area in one of the most important fishing regions in the
world provides a great opportunity to evaluate the behavioral responses
and redistribution of fishing effort by vessels that used to fish there.
\citet{mcdermott_2018} showed that fishing effort within the Phoenix
Islands Marine Protected Area (PIPA) increased between the anouncement
(September 1\textsuperscript{st}, 2014) and its implementation in
Janyary 1\textsuperscript{st}, 2015. Likewise, they demonstrate that
fishing effort is effectively reduced after implementation. To this, we
pose two questions: How do individual vessels respond to the sudden
exclusion of such a big area? And where did all the vessels go? In the
next sections we describe the data and methods used to answer this
questions.

@@

Automatic Identification Systems are on-board devices intended to
provide at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activities to surrounding vessels. These
broadcasted messages can be received by satellites and land-based
antennas. GFW uses a neural network to infer vessel characteristics and
whether each broadcasted position represents a fishing event, thus
allowing us to estimate near real-time fishing events globally since
2012 \citep{kroodsma_2018}. Our data contain information for 2012 -
2017. The recent addition of satellites that can receive AIS signals
causes an apparent increase in the number of broadcasted AIS messages
(\emph{i.e.} points), and therefore number of vessels and fishing hours.
The variability in AIS data and ocean conditions require that temporal
trends be taken into account. We do that by obtaining a subset of data
that meet a BACI design, which gives us the full tracks for vessels
affected and unaffected by the implementation of PIPA.

@@

The treatment group contains all purse seiners (n = 61) that fished
within PIPA at least once before the anouncement (09/01/2014), and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure. This definition of control vessels provides us with 26 purse
seiners that never fished inside PIPA but fished in other PNA waters
before and after PIPA implementation. During our causal identification,
we also use two more strict definitions of control groups: one with only
vessels that belong to PNA countries (n = 7), and one that excludes
Chinese vessels (n = 21). Our definition of treatment and control groups
leaves us with 61 and 21 treated and control vessels, which have just
over 22 million observations where about 22\% are identified as fishing
events. For each of the remainig vessels, we calculate their total daily
fishing hours and obtain a pannel data with 38,200 observations. Table
\ref{tab:baci_n_s} shows the number of vessels following a BACI design,
as well as the fishing hours, before and after PIPA. Figure
\ref{fig:baci_strict} provides a visual representation of the
vessel-level fishing events that make up each group through time.

@@

@@

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month-, and flag-level dummies that account for seasonality or
country-level management
interventions\footnote{We test for other more complex specifications that interact a quarterly or year-month variable with the treatment group and find qualitativel the same results.}

Our second part of the analyses focuses on the redistribution of fishing
effort. In other words, identifying where do vessels that used to fish
inside PIPA go after its establishment. We discretize spatial units by
using a polygon for PIPA
\footnote{we would expect to see a decrease here} and distinct spatial
units for each EEZ of each country. Some vessels might shift from EEZs
into the High Seas, but we are interested in knowing \emph{where} in the
High Seas, so we incorporate additional regions by using a 1 degree
buffer of the High Seas arround each of the EEZ regions. The rest of the
high seas are merged into a single spatial unit. For example, if we were
to do this only for Kiribati, we would have 8 spatial units: PIPA, three
EEZs, three 1-degree buffers of High Seas arround each EEZs, and the
rest of the High Seas. Whenever the buffers overlapped between
themselves, we randocmply clipped one on to the other. EEZs that had
sporadic fishing events were pooled into a group of ``others''.

@@

\[
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that country \(i\) receives at time \(t\). \(Post\) also
represents a policy dummy that takes the value of 0 for all dates before
implementation of PIPA, and 1 otherwise. \(Country\) is a dummy variable
for countries for the spatial units defined above. Our parameter of
interest is \(\beta_{1,i}\), which captures the country-level change in
proportional fishing effort.

@@

\caption{\label{tab:unnamed-chunk-9}\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.}
\centering
\begin{tabular}[t]{lr}
\toprule
country & change\\
\midrule
EEZ COK 1 & 0.53\\
EEZ FSM 1 & 0.81\\
EEZ KIR 1 & -0.16\\
EEZ KIR 2 & 4.50\\
EEZ KIR 3 & -2.79\\
\addlinespace
EEZ MHL 1 & -0.44\\
EEZ NRU 1 & 0.07\\
EEZ PNG 2 & -9.07\\
EEZ SLB 1 & 3.18\\
EEZ TUV 1 & 1.27\\
\addlinespace
HS & 3.93\\
HS COK 1 & 0.06\\
HS FSM 1 & 0.01\\
HS KIR 1 & 2.96\\
HS KIR 2 & 1.07\\
\addlinespace
HS KIR 3 & 2.53\\
HS MHL 1 & 0.11\\
HS MUS 1 & 0.00\\
HS NRU 1 & -0.08\\
HS PNG 2 & -0.20\\
\addlinespace
HS SLB 1 & -0.02\\
HS TUV 1 & 0.01\\
PIPA PIPA 1 & -8.25\\
\bottomrule
\end{tabular}
\end{table}

@@

Our data suggest that purse seiners and longliners have different
responses to the implementation of a Large-Scale Marine Protected Areas.
Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse
seiners have an abrupt increase, just before January 1st, 2015. This
trend is observed for both treated and control groups. The pattern
closely corresponds with the increase in total hours, but the total
number of vessels doesn't entirely follow this pattern. The increase in
fishing hours might be caused by the increased number of
satellites\footnote{Need to check this. Not sure any satellites were
  incorporated during 2014. It is also possible that PNA countries
  started enforcing the requirement of having an AIS unit. On either
  case, both treatment and control groups seem to be affected equally.}.
Longliners, however, show no apparent trend with a clear seasonality
\citep{ortuocrespo_2018}. The number of mmsi codes also increases
slightly through time, but becomes stable after 2015. For both gears and
across all measures, the treatment and control vessels follow similar
patterns, confirming our claim that the control group provides a
plausible counterfactual.

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{\label{fig:unnamed-chunk-10}\label{fig:all_vessels}Fishing hours
and number of vessels by month for all vessels. Vertical dashed line
indicates PIPA closure.}
\end{figure}

@@

>>>>

  \title{Displacement of fishing effort by Large Scale Marine Protected
Areas\footnote{Work in progress, do not circulate}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Updated on 2018-10-18}
  \author{Juan Carlos Villaseñor-Derbez\textsuperscript{1} John
Lynham\textsuperscript{2}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{\textsuperscript{1}Bren School of Environmental Science and Management,
University of California Santa Barbara, Santa Barbara,
CA\newline \textsuperscript{2}Department of Economics, University of
Hawaii at Manoa, Honolulu, HI}

\usepackage{float}
\floatplacement{figure}{H}
\usepackage{lineno}
\linenumbers
\usepackage[bottom]{footmisc}
\usepackage{pdflscape}
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

@@

Marine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environments by
2020. In an effort to meet this target, the world has seen a rapid
increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by
a small number of Large Scale Marine Protected Areas (LSMPA)
\citet{singleton_2014,boonzaier_2016,alger_2017}). Due to weak property
rights, limited habitat transformation, and potentially lower management
costs, LSMPAs provide an opportunity to safeguard the oceans
\citet{game_2009}. Today, a small number of LSMPAs make up at least 80\%
of the managed areas in the ocean \citet{toonen_2013}.

Given the relatively recent establishment of most LSMPAs, very little is
known about their human dimensions and implication for fisheries
\citep{gray_2017}. As with customary MPAs, it is important that we
understand the socioeconomic implications of management interventions.
One issue of particular importance is that of the displacement or
redistribution of fishing effort, which may influence the outcomes of an
MPA \citep{smith_2003}. Theoretical models make different assumptions
about the way in which fishers will reallocate fishing effort after an
area closure \citep{}. The few empirical works focus on small-scale
fisheries and rarely incorporate spatially explicit data. Despite these
efforts, marine conservation is yet to identify the implications of
these management interventions.

Here, we ask, how do fishers respond to the implementation of LSMPA?
After a closure, and if they continue to fish, where do they reallocate
fishing effort? We use novel satellite technologies and causal inference
techniques to identify the behavioral changes and spatial redistribution
of the industrial tuna purse seine fleet due to the implementation of a
Large Scale Marine Protected Areas in the Western and Central Pacific
Ocean (WCPO).

The cutoff at which an MPA is considered to be a LSMPA ranges from areas
larger than 30,000 km\textsuperscript{2} as defined by
\citet{desanto_2013} or areas larger than 250,000 km\textsuperscript{2},
as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows
LSMPAs that meet the later, and that are fully no take. LSMPAs are often
implemented in the pelagic environment, where the dominant human
activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The
early literature on LSMPAs focused on the inherent challenges and
difficulties that come with a pelagic environment. \citet{kaplan_2010}
claimed that very large MPAs would result in excessive opportunity costs
and that these would be difficult to enforce. \citet{game_2009}
suggested that most of the challenges could be overcome with the
incorporation of technology, in what then became known as Dynamic Ocean
Management \citep{maxwell_2015}. The effectiveness and difficulties of
implementating LSMPAs are not the focus of this paper
\footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}.
Instead, we focus on the effect and implications of the ones that
\emph{have been} implemented.

@@

LSMPAs were erroneously assumed to have little social implications due
to their remoteness. However, there have been calls to incorporate the
human dimensions into LSMPAs management and evaluation
\citep{agardy_2011,gray_2017}. Most research incorporating these
dimensions has focused on governance and enforcement of LSMPAs
(\emph{i.e.} \citet{alger_2017,christie_2017}), but they are yet to be
the focus of economic analyses \citep{gray_2017}. Overall, there has
been little empirical work regarding LSMPAs. Recent technological
advances in vessel-detection systems allows for the discovery and
advancement of many important facets of LSMPAs. For example,
\citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to
preemptive overfishing, which can erode or delay the expected benefits
of the intervention. \citet{white_2017} combine shark tags and
vessel-tracking data to demonstrate that the fairly large Palmyra Atoll
National Wildlife Refuge (54,000 Km\textsuperscript{2}) protectes two
thrids of the tagged grey reef sharks by effectively excluding fishing
effort. More recently, \citep{bradley_2018} use similar data to
highlight cases of illegal shark fishing \emph{inside} a 2 million
km\textsuperscript{2} shark sanctuary. To date, no studies have
evaluated the displacement of fishing effort due to LSMPA
implementation.

The exclusion of these LSMPAs is likely to change fisher's behavior.
Theoretical models of fishing effort redistribution range from the
simplistic assumption that effort inside the bounded region disappears,
to spatially explicit models that reallocate fishing effort based on
habitat characteristics, presence of other vessels, and expected returns
\citep{smith_2003,hilborn_2006}. However, these focus on the long term
optimal equilibrium and redistribution of fishing effort may not be
optimal, especially over the first years \citep{stevenson_2013}. The
empirical research that has been done in custumary sized MPAs suggest
that resource users may show idiosyncratic responses. For example,
\citet{stevenson_2013} show that a network of MPAs displaced fishing
effort farther away from ports, resulting in higher \emph{perceived}
costs, and increases in catch per unit effort. \citet{cabral_2017}
analyse the redistribution of fishing and non-fishing vessels following
the implementation of a network of MPAs in California, and find that
commercial dive boats follow a fishing-the-line pattern, while some
fishing boats follow an ideal free distribution. The way in which
fishers react to a spatial closure can have major implications in its
outcome \citep{smith_2003,hilborn_2006} highlighting the need to
understand how fishers react to the implementation of LSMPAs, and
fishing effort changes and is spatially redistributed.

Our work is novel in the sense that it provides empirical evidence of
the effect of large-acale Marine Protected Areas in fishing behavior and
distribution and can help guide future interventions as countries march
towards meetings the targeted 10\% of protection. Furthermore, a growing
body of literature suggests that closing the high seas to all fishing
could increase fishery yields and profitability of fisheries, with
negligible costs to food security
\citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}. Understanding
how effort is displaced from LSMPAs might provide insights into how the
industrial fleet would react to a high seas closure.

The paper is outlined as follows: Section \ref{background} provides an
overview of the Nauru Agreement and associated countries, a description
of the fleet that operates in the region, and a brief history of PIPA.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, section \ref{discussion}
provides an extension of our results to other
LSMPAs\footnote{Not yet. but I think this would be interesting} and
discusses our
results\footnote{I might also include a section, after the background, where I explain the different theoretical models of effort redistribution and behavioral responses}.

@@

The Nauru Agreement was established in 1982 by a select group of Pacific
island nations to manage their important tuna resources. PNA Members
include Federated States of Micronesia, Kiribati, Marshall Islands,
Nauru, Palau, Papua New Guinea, Solomon Islands, and Tuvalu. The Nauru
Agreement regulated access of foreign vessels (\emph{i.e.} those from
non-PNA countries). Holding \textasciitilde{}80\% of the historical
purse seining grounds within their Exclusive Economic Zones, PNA
countries gained bargaining power when providing access to foreign
fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent
agreements that strengthened fisheries management, like the Palau
Agreement, which limited the number of purse seiners at 205 vessels from
1995-2007\footnote{See \citet{havice_2010} for a detailed description of
  the Nauru, Palau, and Federal States of Micronesia agreements, their
  objectives, and outcomes.}. However, the most notable regulation is
their approach to manage fishing effort: a Vessel Day Scheme (VDS)
implemented in 2007 \citep{havice_2013}. This effectively modified how
fishing effort was managed, from number of vessels under the Palau
Agreement to fishing hours. The VDS works as follows: Each year,
scientific advisors recomend a total number of fishing vessel-days per
year. Hours are allocated to each PNA country based on catch history,
and they then use or sell fishing rights to other non-PNA countries
\citep{aqorau_2018}. While the effectiveness of this scheme has been
debated in terms of meeting their fishery management and conservation
objectives, the licensing significantly contributes to the economy of
these island nations \citep{havice_2010}.

@@

The main tuna species caught in the region are skipjack
(\emph{Katsuwonus pelamis}), yellowfin (\emph{Thunnus albacares}),
albacore (\emph{Thunnus alalunga}) and bigeye (\emph{Thunnus obesus}).
From these, the first two are amongst the top-10 species represented in
global fisheries production statistics, with 2016 catches increasing
relative to the 2005-2014 average \citep{fao_2018}. Theis region of the
Pacific has historically accounted for a large portion of tuna catches
\citep{aqorau_1997}. Today, the PNA controls close to 50\% of the global
skipjack tuna production \citep{pna_website_2018}. A large portion of
these catches derive from purse seine vessels licensed under the VDS.
Fishing vessels from Australia, New Zeland, China, France, Korea, Japan,
the Philippines, Taiwan, and the United States participate in the
purse-seining VDS.

One of the most notable and recent management interventions in the
region is the implementation of the Phoenix Island Protected Area (PIPA)
by the government of Kiribati. PIPA was first declared in 2006, and
established in 2008 with only 4\% of it was declared as no-take. In
January 1\textsuperscript{st}, 2015, the no-take area within PIPA was
expanded to a total area of 397,447 km\textsuperscript{2}, roughly 1.5
times the size of Ecuador. Figure \ref{fig:PNA_map} shows a map of the
PNA countries and the Phoenix Island Protected Area.

The implementation of such a large conservation area in one of the most
important fishing regions in the world provides a great opportunity to
evaluate the behavioral responses and redistribution of fishing effort
by vessels that used to fish there. \citet{mcdermott_2018} showed that
fishing effort within the Phoenix Islands Marine Protected Area (PIPA)
increased between the anouncement (September 1\textsuperscript{st},
2014) and its implementation in Janyary 1\textsuperscript{st}, 2015.
Likewise, they demonstrate that fishing effort is effectively reduced
after implementation. To this, we pose two questions: How do individual
vessels respond to the sudden exclusion of such a big area? And where
did all the vessels go? In the next sections we describe the data and
methods used to answer this questions.

@@

Automatic Identification Systems are on-board devices intended to
provide at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activities to surrounding vessels. These
broadcasted messages can be received by satellites and land-based
antennas. GFW uses convolutional neural networks to infer vessel
characteristics and whether each broadcasted position represents a
fishing event, thus allowing us to estimate near real-time fishing
events globally since 2012 \citep{kroodsma_2018}. Our data contain
information for 2012 - 2017. The amount of data gathered is dependent on
the number of antenas and satellites that can pick up the signals. In
June 1\textsuperscript{st} 2014 satellite count increased from 3 to 6,
and then to 10 in January 1\textsuperscript{st} 2016. This causes an
apparent increase in the number of broadcasted AIS messages (\emph{i.e.}
points), and therefore number of vessels and fishing hours. The
variability in AIS data and ocean conditions require that temporal
trends be taken into account. We do that by including specific controls
in our identification strategy and using a subset of data that meet a
BACI design -which gives us the full tracks for vessels affected and
unaffected by the implementation of PIPA.

@@

The treatment group contains all purse seiners (n = 61) that fished
within PIPA at least once before the anouncement (09/01/2014), and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
vessels never fished within PIPA waters, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure. This definition of control vessels provides us with 26 purse
seiners that never fished inside PIPA but fished in other PNA waters
before and after PIPA implementation. During our causal identification,
we also use two more strict definitions of control groups: one with only
vessels that belong to PNA countries (n = 7), and one that excludes
Chinese vessels (n = 21). Our definition of treatment and control groups
leaves us with 61 and 21 treated and control vessels, which have just
over 22 million observations where about 22\% are identified as fishing
events. For each of the remainig vessels, we calculate their total daily
fishing hours and obtain pannel data with 37,800 observations. Table
\ref{tab:baci_n_s} shows the number of vessels following a BACI design,
as well as the fishing hours, before and after PIPA. Fig.
\ref{fig:all_vessels} shows that mean fishing hours for purse seiners
have an abrupt increase, just before January 1st, 2015. This trend is
observed for both treated and control groups. Across all measures, the
treatment and control vessels follow similar patterns, confirming our
claim that the control group provides a plausible counterfactual. Figure
\ref{fig:baci_strict} provides a visual representation of the
vessel-level fishing events that make up each group through time.

@@

\begin{figure}
\centering
\includegraphics{C:/Users/JC/Documents/GitHub/MPA_displacement/docs/Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:unnamed-chunk-7}\label{fig:all_vessels}Fishing hours and
number of vessels by month for all vessels. Vertical dashed lines
indicates dates when satellites were added, solid line indicates PIPA
closurePIPA closure.}
\end{figure}

@@

\[
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \theta_i + \epsilon_{i,t}
\]

Where \(y_{i,t}\) is the variable of interest for vessel \(i\) in time
period \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
including and following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the control
(\(Treat_i = 0\)) or treatment (\(Treat_i = 1\)) group. \(\alpha\) is
the standard intercept, \(\beta_1\) captures the temporal trend change,
\(\beta_2\) captures the difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: de DiD estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month-, and flag-level dummies that account for seasonality or
country-level management interventions, and \(\theta_i\) are a set of
dummies that account for the addition of more
satellites\footnote{We test for other more complex specifications that interact a quarterly or year-month variable with the treatment group and find qualitativel the same results.}.

Our second part of the analyses focuses on the redistribution of fishing
effort. In other words, identifying where do vessels that used to fish
inside PIPA go after its establishment. We discretize spatial units by
using a polygon for PIPA
\footnote{we would expect to see a decrease here} and distinct spatial
units for each EEZ of each country. Some vessels might shift from EEZs
into the high seas, but we are interested in knowing \emph{where} in the
high seas, so we incorporate additional regions by using a 1 degree
buffer of the high seas arround each of the EEZ regions. The rest of the
high seas are merged into a single spatial unit. For example, if we were
to do this only for Kiribati, we would have 8 spatial units: PIPA, three
EEZs, three 1-degree buffers of high seas arround each EEZs, and the
rest of the high seas. Whenever the buffers overlapped between
themselves, we randocmply clipped one on to the other. EEZs that had
sporadic fishing events were pooled into a group of ``others''.

@@

\[
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \theta_i + \epsilon_{i,t}
\]

Our variable of interest, \(y_{i,t}\) represents the proportion of
fishing hours that country \(i\) receives at time \(t\). \(Post\) also
represents a policy dummy that takes the value of 0 for all dates before
implementation of PIPA, and 1 otherwise. \(Country\) is a dummy variable
for countries for the spatial units defined above. Our parameter of
interest is \(\beta_{1,i}\), which captures the country-level change in
proportional fishing effort. \(\theta_i\) accounts for the addition of
satellites.

@@

\caption{\label{tab:unnamed-chunk-10}\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.}
\centering
\begin{tabular}[t]{lr}
\toprule
country & change\\
\midrule
EEZ COK 1 & 0.53\\
EEZ FSM 1 & 0.81\\
EEZ KIR 1 & -0.16\\
EEZ KIR 2 & 4.50\\
EEZ KIR 3 & -2.79\\
\addlinespace
EEZ MHL 1 & -0.44\\
EEZ NRU 1 & 0.07\\
EEZ PNG 2 & -9.07\\
EEZ SLB 1 & 3.18\\
EEZ TUV 1 & 1.27\\
\addlinespace
HS & 3.93\\
HS COK 1 & 0.06\\
HS KIR 1 & 2.96\\
HS KIR 2 & 1.07\\
HS KIR 3 & 2.53\\
PIPA PIPA 1 & -8.25\\
\bottomrule
\end{tabular}
\end{table}

@@

@@

\clearpage
\begin{landscape}
